Datasets used in thesis

### 2014-12-23 Rmarkdown/datasets.Rmd
Figure fixing.
Typo

1. Genetic variants were called using `Bis-SNP (v0.82.2)` via the `bissnp_easy_usage.pl` script.
2. M-bias was estimated using `bismark_methylation_extractor` with the `--mbias_only` flag set. These output files were then analysed using the  `MethylationTuples` R package (see section \ref{sec:MethylationTuples}) and all read-positions with an CpG normalised read-position methylation level more than $0.03$ from the median, i.e., $|$nrplml_{CpG}$ - median(rplm_{CpG})| > 0.03$, were excluded from methylation calling (read 1 and read 2 analysed separately where applicable).
3. CpG methylation calling was performed using `methtuple (v1.4.0)` for m-tuples $\text{m} = 1, \ldots, 8$. CpG 2-tuples were called both with and without the `--all-combinations` flag; all other tuple sizes were called without the `--all-combinations` flag.
4. A `MethPat` object was created for each of the EPISCOPE, Lister, Seisenberger and Ziller datasets for 1-tuples, 2-tuples, 2ac-tuples (2-tuples with `--all-combinations`), 3-tuples and 4-tuples. I did not create `MethPat` objects for m-tuples with $\text{m} > 4$ because the data are too sparse at this larger sizes to be generally useful.

@@

\begin{table}[h!]
\centering
\begin{tabulary}{\textwidth}{LLLLLL}
\toprule
Sample              & Tissue type             & Sequencing & Read length & Ave. coverage & First published             \\ \midrule
ADS                 & Adipose                 & PE         & $75$ bp     & $23\times$    & \citet{Lister:2011kg}       \\
ADS-adipose         & Adipocytes from ADS     & PE         & $75$ bp     & $24\times$    & \citet{Lister:2011kg}       \\
ADS-iPSC            & iPSC from ADS           & PE         & $75$ bp     & $26\times$    & \citet{Lister:2011kg}       \\
FF                  & Foreskin fibroblasts    & SE         & $85$ bp     & $16\times$    & \citet{Lister:2011kg}       \\
FF-iPSC\_6.9        & iPSC from FF            & SE         & $85$ bp     & $10\times$    & \citet{Lister:2011kg}       \\
FF-iPSC\_19.7       & iPSC from FF            & SE         & $85$ bp     & $9\times$     & \citet{Lister:2011kg}       \\
FF-iPSC\_19.11      & iPSC from FF            & SE         & $85$ bp     & $8\times$     & \citet{Lister:2011kg}       \\
FF-iPSC\_19.11+BMP4 & IVD from FF-iPSC\_19.11 & SE         & $85$ bp     & $17\times$    & \citet{Lister:2011kg}       \\
IMR90\_r1           & Lung fibroblasts        & SE         & $87$ bp     & $14\times$    & \citet{Lister:2009hy}       \\
IMR90\_r2           & Lung fibroblasts        & SE         & $87$ bp     & $15\times$    & \citet{Lister:2009hy}       \\
IMR90-iPSC          & iPSC from IMR90         & SE         & $85$ bp     & $9\times$     & \citet{Lister:2011kg}       \\
H1\_r1              & ESC                     & SE         & $85$ bp     & $15\times$    & \citet{Lister:2009hy}       \\
H1\_r2              & ESC                     & SE         & $85$ bp     & $14\times$    & \citet{Lister:2009hy}       \\
H1+BMP4             & IVD from H1             & SE         & $85$ bp     & $33\times$    & \citet{Lister:2011kg}       \\
H9                  & ESC                     & SE         & $85$ bp     & $9\times$     & \citet{Lister:2011kg}       \\
H9\_Laurent         & ESC                     & PE         & $75$ bp     & $8\times$     & \citet{Laurent:2010iw}      \\
HSF1                & ESC                     & SE         & $47$        & $5\times$     & \citet{Chodavarapu:2010iqa} \\ \bottomrule
\end{tabulary}
\caption{Samples in Lister dataset. The reported read lengths are prior to any trimming of the reads. Abbreviations: \em{iPSC} = induced pluripotent stem cell; \em{ESC} = embryonic stem cell; \em{IVD} = \em{in vitro} differentiated from pluripotent cell line \em{PE} = single-end sequencing; \em{PE} = paired-end sequencing.}
\label{tab:Lister_dataset}
\end{table}

@@

\begin{table}[h!]
\centering
\begin{tabulary}{\textwidth}{LLLL}
\toprule
Sample & Sequencing & Read length & Ave. coverage \\ \midrule
E13BUF & PE         & $101$ bp    & $8\times$     \\
E13SA  & PE         & $101$ bp    & $28\times$    \\
E13VA  & PE         & $150$ bp    & $27\times$    \\
E13VAT & PE         & $101$ bp    & $25\times$    \\
E18BUF & PE         & $101$ bp    & $21\times$    \\
E18SA  & PE         & $101$ bp    & $25\times$    \\
E18VA  & PE         & $150$ bp    & $36\times$    \\
E18VAT & PE         & $101$ bp    & $26\times$    \\
E23BUF & PE         & $101$ bp    & $12\times$    \\
E23SA  & PE         & $101$ bp    & $29\times$    \\
E23VA  & PE         & $101$ bp    & $32\times$    \\
E23VAT & PE         & $101$ bp    & $31\times$    \\ \bottomrule
\end{tabulary}
\caption{Samples in EPISCOPE dataset. The reported read lengths are prior to any trimming of the reads.}
\label{tab:EPISCOPE_dataset}
\end{table}

@@

\begin{table}[h!]
\centering
\begin{tabulary}{\textwidth}{LLLLL}
\toprule
Sample            & Tissue                     & Sequencing & Read length & Ave. coverage             \\ \midrule
J1\_1             & ESC                        & PE         & $100$ bp    & $12\times$     \\
E6.5\_epiblast\_1 & Epiblast                   & PE         & $101$ bp    & $13\times$     \\
E16.5\_male\_1    & Male progenitor germ cells & PE         & $150$ bp    & $12\times$     \\ \bottomrule
\end{tabulary}
\caption{Samples in Seisenberger dataset. The reported read lengths are prior to any trimming of the reads. All samples were first published in \citet{Seisenberger:2012ko}.}
\label{tab:Seisenberger_dataset}
\end{table}

@@

\begin{table}[h!]
\centering
\begin{tabulary}{\textwidth}{LLLLL}
\toprule
Sample    & Tissue                     & Sequencing & Read length & Ave. coverage        \\ \midrule
SRR949193 & Frontal\_cortex\_normal\_1 & PE         & 101 bp         & $10\times$ \\
SRR949194 & Frontal\_cortex\_normal\_1 & PE         & 101 bp         & $10\times$    \\
SRR949195 & Frontal\_cortex\_normal\_1 & PE         & 101 bp         & $10\times$    \\
SRR949196 & Frontal\_cortex\_normal\_2 & PE         & 101 bp         & $9\times$     \\
SRR949197 & Frontal\_cortex\_normal\_2 & PE         & 101 bp         & $9\times$     \\
SRR949198 & Frontal\_cortex\_normal\_2 & PE         & 101 bp         & $9\times$     \\
SRR949199 & Frontal\_cortex\_AD\_1     & PE         & 101 bp         & $9\times$     \\
SRR949201 & Frontal\_cortex\_AD\_1     & PE         & 101 bp         & $9\times$     \\
SRR949202 & Frontal\_cortex\_AD\_2     & PE         & 101 bp         & $10\times$    \\
SRR949203 & Frontal\_cortex\_AD\_2     & PE         & 101 bp         & $10\times$    \\
SRR949206 & HepG2\_cell\_line          & PE         & 101 bp         & $2\times$     \\
SRR949207 & HepG2\_cell\_line          & PE         & 101 bp         & $1\times$     \\
SRR949208 & IMR90\_cell\_line          & PE         & 101 bp         & $1\times$     \\
SRR949209 & IMR90\_cell\_line          & PE         & 101 bp         & $3\times$     \\
SRR949210 & Colon\_Tumor\_Primary      & PE         & 101 bp         & $8\times$     \\
SRR949211 & Colon\_Tumor\_Primary      & PE         & 101 bp         & $8\times$     \\
SRR949212 & Colon\_Tumor\_Primary      & PE         & 101 bp         & $9\times$     \\
SRR949213 & Colon\_Tumor\_Primary      & PE         & 101 bp        & $8\times$     \\
SRR949215 & Colon\_Primary\_Normal     & PE         & 101 bp         & $8\times$     \\ \bottomrule
\end{tabulary}
\caption{Sequencing runs in Ziller dataset. The reported read lengths are prior to any trimming of the reads. All samples were first published in \citet{Ziller:2013fs}.}
\label{tab:Ziller_dataset_sequencing_runs}
\end{table}

>>>>

1. Genetic variants were called using `Bis-SNP (v0.82.2)` via the `bissnp_easy_usage.pl` script.
2. M-bias was estimated using `bismark_methylation_extractor` with the `--mbias_only` flag set. These output files were then analysed using the  `MethylationTuples` R package (see section \ref{sec:MethylationTuples}) and all read-positions with an CpG normalised read-position methylation level more than $0.03$ from the median, i.e., $|nrplml_{CpG} - median(rplm_{CpG})| > 0.03$, were excluded from methylation calling (read 1 and read 2 analysed separately where applicable).
3. CpG methylation calling was performed using `methtuple (v1.4.0)` for m-tuples $\text{m} = 1, \ldots, 8$. CpG 2-tuples were called both with and without the `--all-combinations` flag; all other tuple sizes were called without the `--all-combinations` flag.
4. A `MethPat` object was created for each of the EPISCOPE, Lister, Seisenberger and Ziller datasets for 1-tuples, 2-tuples, 2ac-tuples (2-tuples with `--all-combinations`), 3-tuples and 4-tuples. I did not create `MethPat` objects for m-tuples with $\text{m} > 4$ because the data are too sparse at this larger sizes to be generally useful.

@@

\begin{table}[!htb]
\centering
\begin{tabulary}{\textwidth}{LLLLLL}
\toprule
Sample              & Tissue type             & Sequencing & Read length & Ave. coverage & First published             \\ \midrule
ADS                 & Adipose                 & PE         & $75$ bp     & $23\times$    & \citet{Lister:2011kg}       \\
ADS-adipose         & Adipocytes from ADS     & PE         & $75$ bp     & $24\times$    & \citet{Lister:2011kg}       \\
ADS-iPSC            & iPSC from ADS           & PE         & $75$ bp     & $26\times$    & \citet{Lister:2011kg}       \\
FF                  & Foreskin fibroblasts    & SE         & $85$ bp     & $16\times$    & \citet{Lister:2011kg}       \\
FF-iPSC\_6.9        & iPSC from FF            & SE         & $85$ bp     & $10\times$    & \citet{Lister:2011kg}       \\
FF-iPSC\_19.7       & iPSC from FF            & SE         & $85$ bp     & $9\times$     & \citet{Lister:2011kg}       \\
FF-iPSC\_19.11      & iPSC from FF            & SE         & $85$ bp     & $8\times$     & \citet{Lister:2011kg}       \\
FF-iPSC\_19.11+BMP4 & IVD from FF-iPSC\_19.11 & SE         & $85$ bp     & $17\times$    & \citet{Lister:2011kg}       \\
IMR90\_r1           & Lung fibroblasts        & SE         & $87$ bp     & $14\times$    & \citet{Lister:2009hy}       \\
IMR90\_r2           & Lung fibroblasts        & SE         & $87$ bp     & $15\times$    & \citet{Lister:2009hy}       \\
IMR90-iPSC          & iPSC from IMR90         & SE         & $85$ bp     & $9\times$     & \citet{Lister:2011kg}       \\
H1\_r1              & ESC                     & SE         & $85$ bp     & $15\times$    & \citet{Lister:2009hy}       \\
H1\_r2              & ESC                     & SE         & $85$ bp     & $14\times$    & \citet{Lister:2009hy}       \\
H1+BMP4             & IVD from H1             & SE         & $85$ bp     & $33\times$    & \citet{Lister:2011kg}       \\
H9                  & ESC                     & SE         & $85$ bp     & $9\times$     & \citet{Lister:2011kg}       \\
H9\_Laurent         & ESC                     & PE         & $75$ bp     & $8\times$     & \citet{Laurent:2010iw}      \\
HSF1                & ESC                     & SE         & $47$        & $5\times$     & \citet{Chodavarapu:2010iqa} \\ \bottomrule
\end{tabulary}
\caption{Samples in Lister dataset. The reported read lengths are prior to any trimming of the reads. Abbreviations: \em{iPSC} = induced pluripotent stem cell; \em{ESC} = embryonic stem cell; \em{IVD} = \em{in vitro} differentiated from pluripotent cell line \em{PE} = single-end sequencing; \em{PE} = paired-end sequencing.}
\label{tab:Lister_dataset}
\end{table}

@@

\begin{table}[!htb]
\centering
\begin{tabulary}{\textwidth}{LLLL}
\toprule
Sample & Sequencing & Read length & Ave. coverage \\ \midrule
E13BUF & PE         & $101$ bp    & $8\times$     \\
E13SA  & PE         & $101$ bp    & $28\times$    \\
E13VA  & PE         & $150$ bp    & $27\times$    \\
E13VAT & PE         & $101$ bp    & $25\times$    \\
E18BUF & PE         & $101$ bp    & $21\times$    \\
E18SA  & PE         & $101$ bp    & $25\times$    \\
E18VA  & PE         & $150$ bp    & $36\times$    \\
E18VAT & PE         & $101$ bp    & $26\times$    \\
E23BUF & PE         & $101$ bp    & $12\times$    \\
E23SA  & PE         & $101$ bp    & $29\times$    \\
E23VA  & PE         & $101$ bp    & $32\times$    \\
E23VAT & PE         & $101$ bp    & $31\times$    \\ \bottomrule
\end{tabulary}
\caption{Samples in EPISCOPE dataset. The reported read lengths are prior to any trimming of the reads.}
\label{tab:EPISCOPE_dataset}
\end{table}

@@

\begin{table}[!htb]
\centering
\begin{tabulary}{\textwidth}{LLLLL}
\toprule
Sample            & Tissue                     & Sequencing & Read length & Ave. coverage             \\ \midrule
J1\_1             & ESC                        & PE         & $100$ bp    & $12\times$     \\
E6.5\_epiblast\_1 & Epiblast                   & PE         & $101$ bp    & $13\times$     \\
E16.5\_male\_1    & Male progenitor germ cells & PE         & $150$ bp    & $12\times$     \\ \bottomrule
\end{tabulary}
\caption{Samples in Seisenberger dataset. The reported read lengths are prior to any trimming of the reads. All samples were first published in \citet{Seisenberger:2012ko}.}
\label{tab:Seisenberger_dataset}
\end{table}

@@

\begin{table}[!htb]
\centering
\begin{tabulary}{\textwidth}{LLLLL}
\toprule
Sample    & Tissue                     & Sequencing & Read length & Ave. coverage        \\ \midrule
SRR949193 & Frontal\_cortex\_normal\_1 & PE         & 101 bp         & $10\times$ \\
SRR949194 & Frontal\_cortex\_normal\_1 & PE         & 101 bp         & $10\times$    \\
SRR949195 & Frontal\_cortex\_normal\_1 & PE         & 101 bp         & $10\times$    \\
SRR949196 & Frontal\_cortex\_normal\_2 & PE         & 101 bp         & $9\times$     \\
SRR949197 & Frontal\_cortex\_normal\_2 & PE         & 101 bp         & $9\times$     \\
SRR949198 & Frontal\_cortex\_normal\_2 & PE         & 101 bp         & $9\times$     \\
SRR949199 & Frontal\_cortex\_AD\_1     & PE         & 101 bp         & $9\times$     \\
SRR949201 & Frontal\_cortex\_AD\_1     & PE         & 101 bp         & $9\times$     \\
SRR949202 & Frontal\_cortex\_AD\_2     & PE         & 101 bp         & $10\times$    \\
SRR949203 & Frontal\_cortex\_AD\_2     & PE         & 101 bp         & $10\times$    \\
SRR949206 & HepG2\_cell\_line          & PE         & 101 bp         & $2\times$     \\
SRR949207 & HepG2\_cell\_line          & PE         & 101 bp         & $1\times$     \\
SRR949208 & IMR90\_cell\_line          & PE         & 101 bp         & $1\times$     \\
SRR949209 & IMR90\_cell\_line          & PE         & 101 bp         & $3\times$     \\
SRR949210 & Colon\_Tumor\_Primary      & PE         & 101 bp         & $8\times$     \\
SRR949211 & Colon\_Tumor\_Primary      & PE         & 101 bp         & $8\times$     \\
SRR949212 & Colon\_Tumor\_Primary      & PE         & 101 bp         & $9\times$     \\
SRR949213 & Colon\_Tumor\_Primary      & PE         & 101 bp        & $8\times$     \\
SRR949215 & Colon\_Primary\_Normal     & PE         & 101 bp         & $8\times$     \\ \bottomrule
\end{tabulary}
\caption{Sequencing runs in Ziller dataset. The reported read lengths are prior to any trimming of the reads. All samples were first published in \citet{Ziller:2013fs}.}
\label{tab:Ziller_dataset_sequencing_runs}
\end{table}
